Generative LLM oversampling for the class imbalance problem in hate speech detection

### 2023-05-05 paper/paper.tex
fix mean function

%P5: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance
While these advances in resampling have been occurring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and have been applied to fields such as patent claim generation \cite{hsiang} and healthcare education \cite{sallam}. These LLMs have been investigated as a method for oversampling text data to account for the class imbalance problem, with relatively consistent success \cite{edwards} \cite{usuga} \cite{shaikh}, but their use is not widespread and they have not to our knowledge been applied to hate speech detection.

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
None & 0.667 (0\%) & 0.613 (0\%) \\\hline
Random Undersampling & 0.688 (3.1\%) & {\bf 0.638 (4.1\%)} \\\hline
SMOTE-Tomek & 0.667 (0\%) & 0.614 (0.2\%) \\\hline
Random Oversampling & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
AdaSyn & 0.661 (-0.9\%) & 0.608 (-0.8\%) \\\hline
LLM-Augmented & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & 0.312 (0\%) & 0.265 (0\%) \\\hline
Random Undersampling & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
SMOTE-Tomek & 0.314 (0.6\%) & 0.271 (2.3\%) \\\hline
Random Oversampling & 0.369 (18.2\%) & 0.298 (12.4\%) \\\hline
AdaSyn & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
LLM-Augmented & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
Random Undersampling & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
SMOTE-Tomek & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
Random Oversampling & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
AdaSyn & 0.436 (-16.8\%) & 0.286 (-12\%) \\\hline
LLM-Augmented & 0.302 (-42.4\%) &  0.270 (-16.9\%) \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None &  0.382 (0\%) & 0.328 (0\%) \\\hline
Random Undersampling & 0.414 (8.4\%) & 0.339 (3.4\%) \\\hline
SMOTE-Tomek & 0.421 (10.2\%) & 0.366 (11.6\%) \\\hline
Random Oversampling & 0.400 (4.7\%) & 0.358 (9.1\%) \\\hline
AdaSyn & 0.471 (23.3\%) & 0.322 (-1.8\%) \\\hline
LLM-Augmented &  {\bf 0.538 (40.8\%)} & {\bf 0.433 (32.0\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & 0.452 (0\%) & 0.372 (0\%) \\\hline
Random Undersampling & 0.361 (-20.1\%) & 0.323 (-13.2\%) \\\hline
SMOTE-Tomek & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
Random Oversampling & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
AdaSyn & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
LLM-Augmented & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|l|l|}
\hline
Resampling Method & Political & Economic & Religious & Racial \\\hline
None & 0.575 & 0.020 & 0.245 & 0.367 \\\hline
Random Undersampling & 0.601 & 0.023 & 0.205 & 0.362 \\\hline
SMOTE-Tomek & 0.653 & 0.018 & 0.233 & 0.378 \\\hline
Random Oversampling & {\bf 0.661} & 0.015 & 0.227 & 0.392  \\\hline
AdaSyn & 0.588 & 0.003 & 0.243 & 0.376 \\\hline
LLM-Augmented & 0.570 & {\bf 0.111} & {\bf 0.360} & {\bf 0.424} \\\hline
\end{tabular}
\end{nscenter}
In precision and recall, the model is outperformed in political antisemitism (the largest class), and in recall it is slightly outperformed by random undersampling in economic antisemitism, but it outperforms the other models in religious and racial antisemitism on recall and in all classes other than political in precision.

>>>>

%P5: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance
While these advances in resampling have been occurring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and have been applied to fields such as patent claim generation \cite{hsiang} and healthcare education \cite{sallam}. These LLMs have been investigated as a method for oversampling text data to account for the class imbalance problem, with relatively consistent success over not using resampling \cite{edwards} \cite{usuga} \cite{shaikh}. That being said, these studies do not compare the method with other, simpler or more traditional resampling methods. Drummond \& Holt \cite{drummond} argue that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods. This provides motivation, now that the method has shown success in accounting for the class imbalance, for investigating how it compares to existing methods of resampling and under which circumstances it improves relative to those methods.

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
None & 0.667 (0\%) & 0.613 (0\%) \\\hline
Random Undersampling & 0.689 (3.3\%) & {\bf 0.638 (4.1\%)} \\\hline
SMOTE-Tomek & 0.668 (0.1\%) & 0.614 (0.2\%) \\\hline
Random Oversampling & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
AdaSyn & 0.661 (-0.9\%) & 0.608 (0.8\%) \\\hline
LLM-Augmented & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & 0.312 (0\%) & 0.265 (0\%) \\\hline
Random Undersampling & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
SMOTE-Tomek & 0.289 (-7.4\%) & 0.271 (2.3\%) \\\hline
Random Oversampling & 0.369 (18.3\%) & 0.298 (12.5\%) \\\hline
AdaSyn & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
LLM-Augmented & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
Random Undersampling & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
SMOTE-Tomek & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
Random Oversampling & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
AdaSyn & 0.436 (-16.8\%) & 0.286 (-12.0\%) \\\hline
LLM-Augmented & 0.302 (-42.4\%) & 0.270 (-16.9\%) \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & 0.337 (0\%) & 0.328 (0\%) \\\hline
Random Undersampling & 0.414 (22.8\%) & 0.339 (3.4\%) \\\hline
SMOTE-Tomek & 0.420 (24.6\%) & 0.366 (11.6\%) \\\hline
Random Oversampling & 0.400 (18.7\%) & 0.358 (9.1\%) \\\hline
AdaSyn & 0.471 (39.8\%) & 0.322 (-1.8\%) \\\hline
LLM-Augmented & {\bf 0.509 (51.0\%)} & {\bf 0.433 (32.0\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|}
\hline
Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
None & 0.452 (0\%) & 0.372 (0\%) \\\hline
Random Undersampling & 0.361 (-17.9\%) & 0.323 (-13.2\%) \\\hline
SMOTE-Tomek & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
Random Oversampling & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
AdaSyn & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
LLM-Augmented & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
\end{tabular}

@@

\begin{tabular}{|l|l|l|l|l|}
\hline
Resampling Method & Political & Economic & Religious & Racial \\\hline
None & 0.598 & 0.029 & 0.254 & 0.390 \\\hline
Random Undersampling & 0.601 & 0.024 & 0.205 & 0.362 \\\hline
SMOTE-Tomek & 0.653 & 0.018 & 0.234 & 0.378 \\\hline
Random Oversampling & {\bf 0.661} & 0.015 & 0.227 & 0.392 \\\hline
AdaSyn & 0.602 & 0.017 & 0.243 & 0.390 \\\hline
LLM-Augmented & 0.577 &{\bf  0.111} & {\bf 0.367} & {\bf 0.431} \\\hline
\end{tabular}
\end{nscenter}
In precision and recall, the model is outperformed in political antisemitism (the largest class), and in recall it is slightly outperformed by random undersampling in economic antisemitism, but it outperforms the other models in religious and racial antisemitism on recall and in all classes other than political in precision.
